---
title: "homework_4"
author: "Jose Cruz"
date: "9/19/2021"
output:
  word_document: default
  pdf_document: default
---

# Homework 4

## Question 1

The data set prostate(in R package "faraway") is from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy.

Fit a model with lpsa as the response variable (y) and lcavol as the predictor variable (x)and answer the following questions:

### 1. Construct the analysis of variance table (ANOVA) and test for significance of regression;

```{r setup, include=FALSE}

library(faraway)

x_prostate <- prostate$lcavol

y_prostate <- prostate$lpsa


data(prostate,package = "faraway")

fitted_model <- lm(y_prostate~x_prostate,data= prostate)


anova(fitted_model)



```

The anova table above demonstrates that we have a F score of 111.27 and SSR being 69.003 and SSE being 58.915

### 2.Calculate the coefficient of determination (R\^2)

```{r cars}
# r^2 = SSR/SST, sst=ssr+sse
summary(fitted_model)
```

The calculated coefficient of determination (r\^2) is found in our LM function, looking at the Residual standard error we can see we get 0.7875.

### 3. Find a 95% confidence interval on the slope (β1);

```{r}
n <-length(y_prostate)
SSRes <-sum(fitted_model$residuals^2)
MSReS <- SSRes/(n-2)
Sxx <-sum((x_prostate-mean(x_prostate))^2)
se <- sqrt(MSReS/Sxx)
t.critical <-qt(0.975,df=n-2)


LL.beta1 <- coef(fitted_model)[2]-t.critical*se

UL.beta1 <- coef(fitted_model)[2]+t.critical*se

CI.beta1 <- c(LL.beta1,UL.beta1)
names(CI.beta1) <- c("2.5%", "97.5%")
CI.beta1

```

For the 95 percent confidence interval on the slope of beta 1 we get a lower bound of 0.5839402 and a upper bound of 0.8547001 for the confidence interval of beta 1.

### 4. Find a 95% confidence interval on the mean lpsa(Y) when lcavol= 1.35 (X): 

```{r}
x_bar <-mean(x_prostate)


SSRes <-sum(fitted_model$residuals^2)
MSReS <- SSRes/(n-2)

sxx <- sum((x_prostate-x_bar)^2)

x0 <-1.35

beta0 <-fitted_model$coefficients[1]
beta1 <-fitted_model$coefficients[2]
mu_hat <-beta0+beta1*x0 # just makes the formula

lower_limit <- mu_hat-t.critical*sqrt(MSReS*(1/n+(x0-x_bar)^2/sxx))
                 
lower_limit
                                                                                                 
ul <- mu_hat+t.critical*sqrt(MSReS*(1/n+(x0-x_bar)^2/Sxx))

ul

```

When we find the 95 percent confidence interval on the mean of lpsa when lcavol= 1.35 we a lower limit of 1.935894 and a upper limit of 3.020866 .

### 5.Calculate a 95% prediction interval of lps a when lcavol= 1.35.

```{r}
#prediction interval

y_hat <- beta0+beta1*x0

LL <- y_hat-t.critical*sqrt(MSReS*(1+1/n+(x0-x_bar)^2/Sxx))

UL <- y_hat+t.critical*sqrt(MSReS*(1+1/n+(x0-x_bar)^2/Sxx))


LL
UL
```

Our outcome for our prediction interval at 95 percent when lcavol= 1.35 is a lower limit of -0.542 and a upper limit of 5.499

## Problem 2

Hsuie, Ma and Tsai studied the effect of the molar ratio of sebacic acid (the predictor variable) on the intrinsic viscosity of copolyesters (the response variable). The table gives the data.Fit a linear regression model and answer the following questions.

### 1.Find a 95% confidence interval on the slope (β1).

```{r}

#beta 1 95 percent confidence interval

y_Visocity <- c(0.45,0.2,0.34,0.58,0.7,0.57,0.55,0.44)
x_Ratio<- c(1.0,0.9,0.8,0.7,0.6,0.5,0.4,0.3)

visocityModel <- lm(y_Visocity~x_Ratio,data=prostate)

confint(visocityModel,level=0.95)


 
```

Confidence interval for beta 1 slope of between -0.86 and 0.27

### 2. A 95% confidence interval on the mean viscosity when ratio= 0.65:

```{r}
#
x_bar <-mean(x_Ratio)
n <- length(y_Visocity)

SSRes <-sum(visocityModel$residuals^2)
MSReS <- SSRes/(n-2)

sxx <- sum((x_Ratio-x_bar)^2)

x0 <-0.65

beta0 <-visocityModel$coefficients[1]
beta1 <-visocityModel$coefficients[2]
mu_hat <-beta0+beta1*x0 # just makes the formula
t.critical <-qt(0.975,df=n-2)
lower_limit <- mu_hat-t.critical*sqrt(MSReS*(1/n+(x0-x_bar)^2/sxx))
                 
lower_limit
                                                                                                 
ul <- mu_hat+t.critical*sqrt(MSReS*(1/n+(x0-x_bar)^2/Sxx))

ul




```

Calulating the confidence interval at 95 precent for when ratio equals 0.65 we get a lower bound of 0.349 and a upper bound of 0.609

### 3. Find a 95% prediction interval for viscosity when ratio= 0.65.

```{r}

y_hat <- beta0+beta1*x0

LL <- y_hat-t.critical*sqrt(MSReS*(1+1/n+(x0-x_bar)^2/Sxx))

UL <- y_hat+t.critical*sqrt(MSReS*(1+1/n+(x0-x_bar)^2/Sxx))


LL
UL
```

The calculations for a 95 percent prediction interval when ratio equals 0.65 is a lower limit of 0.089 and a upper limit of 0.868.
