---
title: "Homework6"
author: "Jose Cruz"
date: "10/4/2021"
output: word_document
---

# Homework 4

## Problem 1

### A.

Fit a multiple regression model withtasteas the response and the three chemical contents as pre-dictors (Acetic, H2S, and, Lactic). Report the values of the regression coefficients (ˆβ0,ˆβ1,ˆβ2,ˆβ3).

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)

#laoad datasets

library(faraway)
data(cheddar,package = "faraway")


mutiLinearModel <-lm(taste~Acetic+H2S+Lactic,data = cheddar)

summary(mutiLinearModel)



```

The regression coefficient are:

-   b0: -28.8768
-   b1: 0.3277
-   b2: 3.9118
-   b3: 19.6705

### B.

What change in the taste would be expected, if only H2S is increased by 1 unit according to the fitted model?

```{r }

```

If we increased the H2s by 1 unit according to the fitted model the change is taste is expected to increase by 3.9118

### C. Calculate R2 and adjusted R2 for this model.

```{r}
summary(mutiLinearModel)$r.squared
summary(mutiLinearModel)$adj.r.squared
```

The calculated r\^2 and adjusted r\^2 is 0.6517747 and 0.6115948

### D. Construct the analysis-of-variance table and test for significance of regression.

```{r}

anova(mutiLinearModel)

```

| Col1      | Col2 | Col3    | Col4    | Col5    | Col6      | Col7 |
|-----------|------|---------|---------|---------|-----------|------|
| respsonse | df   | sum sq  | mean sq | F value | P value   |      |
| acetic    | 1    | 2314.14 | 2314.14 | 22.5481 | 6.528e-05 |      |
| h2s       | 1    | 2147.02 | 2147.02 | 20.9197 | 0.0001035 |      |
| lactic    | 1    | 533.32  | 533.32  | 5.1964  | 0.0310795 |      |
| resisual  | 26   | 2668.41 | 102.63  |         |           |      |
| total     | 29   |         |         |         |           |      |

## Problem 2

### A. Report your fitted model;

```{r}

data(sat,package = "faraway")


satModel <-lm(total~expend+ratio+salary,data = sat)

summary(satModel)

```

My fitted model is y=1069.234+16.469x1+ 6.330x2-8.823x3

### B. Interpret your estimated ˆβ salary.

```{r}

```

Our estimated beta4 salary is very significant to our model with a p value of 0.667 and a t value of -1.878
