---
title: "stat3110-Homework8"
author: "Jose Cruz"
date: "10/25/2021"
output: word_document
---

# Homework 8

```{r setup, include=FALSE}

library(faraway)
data(sat,package = "faraway")

sModel <-lm(total~expend+ratio+salary+takers,data=sat)
 


summary(sModel)


```

## A. Check the constant variance assumption by using residuals vs fitted value plot;

```{r }

residuals(sModel)


plot(fitted(sModel), residuals(sModel), xlab="Fitted", ylab="Residuals")
plot(sModel)
#abline(h=0, col="red")


```

Preforming a constant Variance test using our residuals and fitted values, we can see that our model show there is nonlinearlity in our model and evidence of nonconstant variance as well.

## B. Check if errors are uncorrelated by plotting successive pairs of residuals and the Durbin-Watson test.

### Successive Pair of Residual Test

```{r}
summary(sModel)
## Plot successive pairs of residuals
n <- length(residuals(sModel))
plot(tail(residuals(sModel), n-1), head(residuals(sModel), n-1),
xlab=expression(hat(epsilon[i])), ylab=expression(hat(epsilon[i+1])))
#abline(h=0, v=0, col=grey(0.75))

#par(mfrow=c(1, 1))

```

In our Successive pair of residual test we find a negative correlation between successive pairs of residuals, indicating a negative serial correlation and shows errors are uncorrelated.

### DW Test

```{r}
library(lmtest)

dwtest(sModel)
```

Using our Durbin-watson test we see that our p value = 0.9459 and Dw = 2.4525. Looking at our p value we see that our model has low evidence of serial correlation.
