---
title: "R Notebook"
output: html_notebook
---

## problem 1

The data set teengamb(in R package “faraway”, you will need to install the package in R) concerns a study of teenage gambling in Britain

### 1. For each variable in the data set, make a numerical summary and comment of any features youfind interesting (range, mean, five number summary statistics).
```{r}
#load library
library("faraway")

#loads the teen gambling dataset in the faraway complete set
data(teengamb,package="faraway")

#summary of each column with the 6 most basic functions of stats
summary(teengamb)[,1:5]

```


commenting on the variables in this dataset I noticed some things:

1. sex seems to be a categorical type of variable due to the min and max being only 2 numbers
2. status seems seems to be more normally distributed  excluding the outlier in the end with a max of 15
3. income seems to almost distributed  if we include that the mean is being affected by the outlier of 15 being the max
4. verbal seems to be mostly clustered around 6-8 mostly with 1 being an outlier in the data
5. gamble is an interesting variable, median on gambling is 6 but mean is 19.3 ans max is 156 so there are some outliers with heavy numbers that is skewing it right.
________________________________________________________________________________________________________________________________________________

### 3.  Make a histogram and describe its distribution for each quantitative variable in the data set.

```{r}

hist(teengamb$income,xlab="income",main = "")
hist(teengamb$status,xlab="status",main = "")
hist(teengamb$verbal,xlab="verbal",main = "")
hist(teengamb$gamble,xlab="gamble",main = "") 

```
- income: distribution on the income seems to be right tail with most income being between 0-5
 
- Status: status frequency seems to be around the same for most 

- verbal: seems that verbal has a left tail distribution with most verbal levels being around 5-8

- gamble: most of gambling variables are around 0-25 with a right tail distribution 





### 2.  Are there categorical variables in the data set, if yes, which ones?
```{r}

#Test for sex (this is a binomial type so this can work) will classify the  note that 0 is female and 1 is male

teengamb$sex <- factor(teengamb$sex)
levels(teengamb$sex) <- c("male","female") #c makes a columns and makes changes
summary(teengamb$sex)


#Looks at first 10 sections of the dataset
head(teengamb,n=10)

```


yes if we reassign the variables 0 and 1 to male and female we can see that each person is being catageroized.





________________________________________________________________________________________________________________________________________

### 4.  Make pairwise scatter plots (all pairwise). Any linear relationships between these variables?
```{r}

pairs(teengamb [,c(2:5)])
require(ggplot2)

ggplot(na.omit(teengamb), aes(x=income,y=gamble))+geom_point()
```
- There are some relationships that can be put together:


- if we look at the income and gambling we can see most lower income people like to mostly participate in gambling.

- Another relationship we can look is that verbal and status seem to have a correlation where if your verbal level is higher your status is higher lossly related but a interesting relation.


## problem 3

The data set prostate(in R package “faraway”) is from a study on 97 men with prostate cancer who were due to receive a radical prostate to my. Fit a linear model with lpsa as the response variable (y) and lcavol as the predictor variable (x) by using the lm function in R and answer the following questions based on your model fitting results:

### 1. Make a scatter plot of the two variables (lpsa and lcavol), the figure needs to be included in your submission;
```{r}
library("faraway")

data(prostate,package = "faraway")

head(prostate) #has 8 variables and 1 categorical variable

summary(prostate)


library(ggplot2)
ggplot(na.omit(prostate), aes(x=lcavol,y=lpsa))+geom_point()
```


### 2.  What is the fitted regression model?
```{r}
#make a bi variate linear regression model

linearRegressionModel=lm(lpsa~lcavol,data = prostate)

summary(linearRegressionModel)

plot(prostate$lcavol,prostate$lpsa, xlab="lcavol", ylab='lpsa',
	main="lcavol vs lpsa", cex=1)

abline(linearRegressionModel, cex=1, col="red")


#fitted regression values
ggplot(na.omit(prostate), aes(x=lcavol,y=lpsa))+
  geom_point(alpha=0.7)+ #observed data
  geom_point(aes(x=lcavol, y=linearRegressionModel$fitted.values), #predicted data
             color='red',alpha=0.5)+
  geom_segment(aes(xend=lcavol, yend=linearRegressionModel$fitted.values),color='red',linetype='dashed',alpha=0.25)
```

- The fitted regression model is y=(1.50730)+(0.71932)xi for lcavol vs lpsa. 

- fitted graphs are above ^



### 3.  What is the estimate of σ2;
```{r}

#summary of linear relationship for the model
summary(linearRegressionModel)


sigmaSquared <-0.7875^2

sigmaSquared

```


- The estimate of σ2 is 0.6201562. We use our stardard error and square the result to get our estimate variance.













Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
